---
output:
  html_document:
    theme: readable
    toc: yes
---
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
options(scipen=1, digits=4, width=80)
library(faraway)
library(readr)
library(caret)
library(klaR)
library(lattice)
```
# CS 498 AML Homework 5
## Question 1 - Music
```{r, message=FALSE}
music <- read_csv('./music.csv')
music["latitude"] <- music["latitude"] + 90
music["longitude"] <- music["longitude"] + 180
```
**(a)** Fit a linear model
Latitude:
```{r}
lat.lm <- lm(latitude ~ . - longitude, data = music)
print(c("The R^2 value is", summary(lat.lm)$r.squared))
#summary(lat.lm)
xyplot(lat.lm$residuals ~ lat.lm$fitted,
      xlab = "Fitted Values",
      ylab = "Residuals",
      main = "Predicting Latitude",
      panel = function(x, y, ...)
      {
        panel.grid(h = -1, v = -1)
        panel.abline(h = 0)
        panel.xyplot(x, y, ...)
      }
)
```

Longiutude:
```{r}
lon.lm <- lm(longitude ~ . - latitude, data = music)
print(c("The R^2 value is", summary(lon.lm)$r.squared))
#summary(lon.lm)
xyplot(lon.lm$residuals ~ lon.lm$fitted,
      xlab = "Fitted Values",
      ylab = "Residuals",
      main = "Predicting Longitude",
      panel = function(x, y, ...)
      {
        panel.grid(h = -1, v = -1)
        panel.abline(h = 0)
        panel.xyplot(x, y, ...)
      }
)
```

**(b)** Box-Cox
Latitude:
```{r}
lat.bc_info = boxcox(lat.lm, lambda=seq(-1,5))
lat.bc_lambda = lat.bc_info$x[which.max(lat.bc_info$y)]
lat.lm_lambda = lm(latitude^lat.bc_lambda ~ . - longitude, data=music)
print(c("The R^2 value is", summary(lat.lm_lambda)$r.squared))
print(c("The Box-Cox optimal lambda is", lat.bc_lambda))
#summary(lat.lm)
xyplot(lat.lm_lambda$residuals ~ lat.lm_lambda$fitted,
      xlab = "Fitted Values",
      ylab = "Residuals",
      main = "Predicting Latitude",
      sub = toString(c("power", lat.bc_lambda)),
      panel = function(x, y, ...)
      {
        panel.grid(h = -1, v = -1)
        panel.abline(h = 0)
        panel.xyplot(x, y, ...)
      }
)
```

The Box-Cox transformation helps tremendously with the lattitude. This is evident because the residuals have dropped from about $30$ to $10^{-7}$


Longitude:
```{r}
lon.bc_info = boxcox(lon.lm)
lon.bc_lambda = lon.bc_info$x[which.max(lon.bc_info$y)]
lon.lm_lambda = lm(longitude^lon.bc_lambda ~ . - latitude, data=music)
print(c("The R^2 value is", summary(lon.lm_lambda)$r.squared))
print(c("The Box-Cox optimal lambda is", lon.bc_lambda))
xyplot(lon.lm_lambda$residuals ~ lon.lm_lambda$fitted,
      xlab = "Fitted Values",
      ylab = "Residuals",
      main = "Predicting Longitude",
      sub = toString(c("power", lon.bc_lambda)),
      panel = function(x, y, ...)
      {
        panel.grid(h = -1, v = -1)
        panel.abline(h = 0)
        panel.xyplot(x, y, ...)
      }
)
```

The Box-Cox transformation does not help becuase the residuals are not really effected. Since the optimal $\lambda$ found by Box-Cox is about $1$, we don't expect this to do a lot.


## Question 2 - Taiwan

## Question 3 - Cancer
